---
title: "Exploring data"
author: "Juan Rocha"
date: "2025-06-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(network)
library(ergm)
library(fs)
```

## Datasets:

Currently there is the following datasets downloaded from SunNet:

```{r}
dir_info("data/raw/")
```

-   LinkAlter: is the data of connections between alters $i,j$ as reported by node $k$ but it does not include the links from $k,i|k,j$.
-   Alter and Ego are node level attributes from the survey, but they differ in detail. Ego info is more rich, all questions answered. Alter info is as perceived from the Ego, not all questions available.
-   `EGO_ALTER_CLEANED_ONLY PEER` contains the full survey.

```{r warning = FALSE, message=FALSE}
## links: 
link_alter <- read_csv("data/raw/Original Datase_linkAlter_All observations.csv")
alter <- read_csv("data/raw/Original Dataset_Alter_All observations.csv")
ego <- read_csv("data/raw/Original Dataset_Ego_All observations.csv")
mat <- read_csv("data/raw/Original dataset_EGO_ALTER_CLEANED_ONLY PEER.csv")
```

First clean up LinkAlter to create a full network:

```{r}
link_alter
```

```{r}
# rename:
names(link_alter) <- names(link_alter) |> 
    str_remove_all("networkcanvas") |> 
    str_replace(pattern = "uu", replacement =  "_") 

link_alter <- link_alter |> 
    rename(link_id = `_id`) 

# net <- link_alter |> # 3410
#     select(source_id, target_id) |> 
#     bind_rows(link_alter |> select(source_id = ego_id, target_id = source_id) |>
#                   unique() ) |> #3950
#     bind_rows(link_alter |> select(source_id = ego_id, target_id) |> unique()) |> # 4490
#     unique() |> # 3975
#     network(directed = TRUE, matrix.type = "edgelist", ignore.eval = TRUE)

df_mat <- mat |> 
    select(source = idencuesta,  target = idal, ego_id = networkcanvasegouuid, alter_id = networkcanvasuuid) 

df_alters <- link_alter |> 
    filter(connectionalter == 1) |> 
    select(ego_id, alter_id = source_id, target_id) |> 
    left_join(df_mat) |> 
    rename(ego_old = ego_id, ego_id = alter_id, alter_id = target_id) |> 
    select(-ego_old)

df_edges <- bind_rows(df_mat, df_alters) |> 
    filter(!is.na(source), !is.na(target)) |> 
    unique()

df_nodes <- df_edges |> 
    pivot_longer(cols = c(source, target), names_to = "role", values_to = 'node_id') |> 
    pivot_longer(cols = c(ego_id, alter_id), names_to = "roles", values_to = "alias") |> 
    select(node_id, alias) |> 
    unique()

net <- network(
    df_edges |> select(source, target) |> unique(), 
    directed = TRUE, matrix.type = "edgelist", ignore.eval = TRUE
)


plot.network(
    net, displayisolates = TRUE, edge.col = alpha("grey50", alpha = 0.4), vertex.col = "orange",
    vertex.border = 0, edge.lwd = 0.01)

```

Extract the node and edge list to create attributes:

```{r}
nodes <- tibble(node = network.vertex.names(net)) |> 
    # id is the id in the network object, node_id is the id in the database (df_nodes). node_id is need to merge with attributes.
    rownames_to_column("id") |> 
    mutate(id = as.numeric(id)) |> rename(node_id = node)
edges <- as.edgelist(net)

edges <- edges |> as_tibble() |> 
    rename(source = V1, target = V2) |> 
    mutate(source = as.numeric(source), target = as.numeric(target))

edges <- edges |>
    left_join(nodes, by = c("source" = "node_id")) |> 
    rename(source_id = node) |> 
    left_join(nodes, by = c("target" = "node_id")) |> 
    rename(target_id = node)
edges
```

Node attributes: `ego` and `altitude` don't exist on the dataset

```{r}
## Egos
 egos <- mat |> 
    select(node_id = idencuesta, shadechanges, starts_with("ferti"), starts_with("village"), egoleader, leader, starts_with("sex"), starts_with("education"), birth, starts_with("training"), nespresso = certifitype_6, starts_with("cliper"), expe, starts_with("area")) |> 
    unique() |> 
    group_by(node_id) |>
    mutate(leader = case_when(any(leader) ~ TRUE, .default = sum(leader))) |> unique() |> 
    mutate(
        shadechanges = shadechanges,
        fert_org = ferti_5 == TRUE | ferti_6 == TRUE,
        village = case_when(
            village_1 == TRUE ~ "1",
            village_2 == TRUE ~ "2",
            village_3 == TRUE ~ "3"),
        sex = case_when(sex_1 == TRUE ~ "woman", sex_2 == TRUE ~ "man"),
        education = case_when(
            education_1 == TRUE ~ "none", 
            education_2 == TRUE ~ "pre-school", 
            education_3 == TRUE ~ "primary", 
            education_4 == TRUE ~ "secondary", 
            education_5 == TRUE ~ "superior"),
        age = 2025 - birth,
        training_shade = training_1, training_organics = training_2,
        area = case_when(
            areafincauni_1 == TRUE ~ areafarm,
            areafincauni_2 == TRUE ~ areafarm/10000,
            areafincauni_3 == TRUE ~ areafarm*0.643),
        ) |> 
    select(node_id, shadechanges, fert_org, village, sex, education, age, training_shade, training_organics, area, egoleader, leader, nespresso:expe )

# alters
mat |> select(node_id = idal, starts_with("shadealter"), starts_with("alterferti"), starts_with("altervillage"), leader, starts_with("sex"), name) 
```

`shadealter_` is a property of the alter

```{r}
## alters
mat |> select(starts_with("shadealter"))
```

For response variables we probably need to code them as categorical or continuous.
Categorical may be more convenient in ergms, but continuous if used as a response variable in a linear model.
